\documentclass[12pt]{article}
\usepackage[margin=1in, headheight=20pt]{geometry}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{eso-pic}
\usepackage{amsthm, amsmath, amssymb}
\usepackage{mathtools}
\usepackage[italicdiff]{physics}
\usepackage{enumitem}
\usepackage{lmodern}
\usepackage{fancyhdr}
\usepackage{pgfornament}

\definecolor{pagecolor}{HTML}{DCE2F0}
\definecolor{textcolor}{HTML}{373D4A}

\definecolor{quasicolor}{HTML}{5E4E9C}
\definecolor{taxicabcolor}{HTML}{4169B0}
\definecolor{euclideancolor}{HTML}{0080C4}
\definecolor{infinitycolor}{HTML}{009EE3}

\pagecolor{pagecolor}
\color{textcolor}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Numerical Optimization}
\fancyhead[R]{\nouppercase{\leftmark}}
\fancyfoot[C]{}

\renewcommand{\headrule}{
  \vspace{-5pt}
  \hbox to \headwidth{
    \leaders\hrule height 0.5pt\hfill
    \hspace{5pt}
    \raisebox{0.20pt}{\pgfornament[width=1cm]{11}}
    \hspace{5pt}
    \leaders\hrule height 0.5pt\hfill
  }
}

\renewcommand{\footrule}{
  \vspace{-12pt}
  \hbox to \headwidth{
    \leaders\hrule height 3.5pt depth -3pt \hfill 
    \hspace{5pt} 
    \thepage 
    \hspace{5pt}
    \leaders\hrule height 3.5pt depth -3pt \hfill
  }
}

\fancypagestyle{plain}{
  \fancyhf{}
  \renewcommand{\headrulewidth}{0pt}
  \renewcommand{\headrule}{} 
  \fancyfoot[C]{}
  \renewcommand{\footrule}{
    \vspace{-12pt}
    \hbox to \headwidth{
      \rule[0.65ex]{0.47\headwidth}{0.5pt}%
      \hfill
      \thepage
      \hfill
      \rule[0.65ex]{0.47\headwidth}{0.5pt}%
    }
  }
}

\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\cl}[1]{\mathcal{#1}}

\newcommand{\p}[1]{\left ( #1 \right )}
\newcommand{\bk}[1]{\left [ #1 \right ]}
\newcommand{\br}[1]{\left \{ #1 \}}
\newcommand{\ab}[1]{\langle #1 \rangle}

\newcommand{\f}[2]{\frac{#1}{#2}}
\newcommand{\nset}{\varnothing}
\newcommand{\oo}{\infty}

\newcommand{\gm}{\gamma}
\newcommand{\de}{\delta}
\newcommand{\De}{\Delta}
\newcommand{\ep}{\varepsilon}
\newcommand{\la}{\lambda}
\newcommand{\si}{\sigma}
\newcommand{\om}{\omega}
\newcommand{\Om}{\Omega}

\newcommand{\imp}{\Rightarrow}
\newcommand{\pmi}{\Leftarrow}
\renewcommand{\iff}{\Leftrightarrow}
\newcommand{\ffi}{\Rightarrow\!\Leftarrow}

\setlist[enumerate]{label=(\alph*)}
\pgfplotsset{compat=newest}

\newtheoremstyle{boldnote}
  {}
  {}
  {\itshape}
  {}
  {\bfseries}
  {.}
  { }
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (\bfseries #3)}}
\theoremstyle{boldnote}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}{Example}

\title{
    \textbf{Numerical Optimization} \\
}
\author{
    Dhyan Laad \\
    \texttt{2024ADPS0875G}
}
\date{}

\begin{document}
\maketitle

\section{Preliminaries}
\subsection{Inner Product Spaces}
\begin{definition}
    Let $V$ be a real vector space. A function $\ab{\cdot, \cdot} : V \times V \to \bb R$ is called a \emph{real inner product} if it satisfies the following three properties for $x, y, z \in V$ and $c \in \bb R$:
    \begin{enumerate}
        \item $\ab{x, y} = \ab{y, x}$,
        \item $\ab{x + z, y} = \ab{x, y} + \ab{z, y}$,
        \item $\ab{cx, y} = c\ab{x, y}$, and
        \item $\ab{x, x} \geq 0$ and $\ab{x, x} = 0 \iff x = 0$.
    \end{enumerate}
\end{definition}

The inner product is the generalization of the dot product on Euclidean vector spaces: for $\vb x = (x_1, x_2, \dots , x_n)$ and $\vb y = (y_1, y_2, \dots , y_n)$ where $x_i, y_i \in \bb R$ for $i \in 1 : n$,
\[\vb x \cdot \vb y = \sum_{i=1}^n x_iy_i.\]
It may also be defined on $\bb C$-spaces, replacing (a) with conjugate symmetry: $\ab{x, y} = \overline{\ab{y, x}}$, and adding conjugate linearity in the second argument: $\ab{x, cy} = \bar{c}\ab{x, y}$.

\begin{definition}
    Let $V$ be a real vector space and $\ab{\cdot, \cdot}$ an inner product. Then, $(V, \ab{\cdot, \cdot})$ is called a \emph{real inner product space}.
\end{definition}

For brevity, we may simply state that $V$ is an inner product space, with the notation for the inner product being implicit, and all future inner product spaces map into $\bb R$ unless stated otherwise.

\subsection{Normed Linear Spaces}
\begin{definition}
    Let $V$ be a real vector space. A function $\norm{\cdot} : V \to \bb R$ is called a \emph{norm} if it satisfies the following properties for $x, y \in V$ and $c \in \bb R$:
    \begin{enumerate}
        \item $\norm{cx} = \abs{c} \norm{x}$,
        \item $\norm{x} \geq 0$ and $\norm{x} = 0 \iff x = 0$, and
        \item $\norm{x + y} \leq \norm{x} + \norm{y}$.
    \end{enumerate}
\end{definition}

This last property is referred to as the \emph{triangle inequality}. The norm assigns a notion of length to vectors, and generalizes the standard formula for the length of a vector in a Euclidean vector space: for $\vb x = (x_1, x_2, \dots , x_n)^\top$ where $x_i \in \bb R$ for $i \in 1 : n$.
\[\norm{\vb x} = \p{\sum_{i=1}^n x_i^2}^{1/2}.\]

\begin{definition}
    Let $V$ be a vector space and $\norm{\cdot}$ a norm. Then $(V, \norm{\cdot})$ is called a \emph{normed linear space}.
\end{definition}

Once again, we may conventionally omit the norm from notation when defining a new normed linear space.

\begin{definition}
    Let $p \geq 1$. The \emph{$p$-norm} (or \emph{$\ell^p$-norm}) of a vector $\vb x = (x_1, x_2, \dots, x_n)^\top$ where $x_i \in \bb R$ for $i \in 1 : n$ is
    \[\norm{\vb x}_p \coloneqq \p{\sum_{i=1}^n \abs{x_i}^p}^{1/p}.\]

    For $p=1$, we get the \emph{taxicab} or \emph{Manhattan} norm, for $p=2$, we get the standard Euclidean norm, and for $p \to \oo$, the $p$-norm approaches the \emph{infinity} or \emph{maximum} norm:
    \[\norm{\vb x}_\oo \coloneqq \max_{i} \abs{x_i}.\]
\end{definition}

For $p \in (0, 1)$, the triangle inequality does not hold, and the resulting functions are called \emph{quasinorms}. Pictured below are unit circles in the $0.5$-quasinorm, and $p$-norms for $p \in \{1, 2, \oo\}$.

\begin{center}
\begin{tikzpicture}[scale=2.5]

    % Axes
    \draw[->] (-1.5,0) -- (1.5,0) node[right] {$x$};
    \draw[->] (0,-1.5) -- (0,1.5) node[above] {$y$};
    \node[below left, font=\scriptsize] at (0,0) {0};

    % Infinity norm
    \draw[infinitycolor, thick, dashdotted] (-1,-1) rectangle (1,1);
    \node[infinitycolor, above right] at (1,1) {$\oo$};

    % Euclidean norm
    \draw[euclideancolor, thick] (0,0) circle (1cm);
    \node[euclideancolor, above right] at (0.71, 0.71) {2};

    % Taxicab norm
    \draw[taxicabcolor, thick, dashed] (1,0) -- (0,1) -- (-1,0) -- (0,-1) -- cycle;
    \node[taxicabcolor, right] at (0.5, 0.6) {1};

    % Quasinorm (1/2)
    \draw[quasicolor, thick, dotted] 
        plot[domain=0:360, samples=200, smooth] 
        ({sign(cos(\x))*abs(cos(\x))^4}, {sign(sin(\x))*abs(sin(\x))^4});
    \node[quasicolor, below left] at (0.5, 0.5) {$0.5$};

\end{tikzpicture}
\end{center}

\begin{lemma}
  Let $V$ be a normed linear space. Then,
  \[\abs{\norm{x} - \norm{y}} \leq \norm{x - y}\]
  for every $x, y \in V$.
\end{lemma}

\begin{theorem}
  Let $(V, \norm{\cdot})$ be a normed linear space. Then $\norm{\cdot} : V \to \bb R$ is uniformly continuous.
\end{theorem}
\begin{proof}
  Let $\ep > 0$, and choose $\de = \ep$. If $\norm{x - y} < \de$, then by Lemma 1.6,
  \[\abs{\norm{x} - \norm{y}} \leq \norm{x - y} < \de = \ep.\]
  Therefore, $\abs{\norm{x} - \norm{y}} < \ep$.
\end{proof}

\begin{theorem}
  Every inner product space $V$ is naturally a normed linear space. For $x \in V$ define
  \[\norm{x} \coloneqq \sqrt{\ab{x, x}}.\]
\end{theorem}

However, the converse is not always true.

\begin{theorem}
  A normed linear space $V$ is an inner product space iff the norm satisfies the parallelogram law:
  \[\norm{x + y}^2 + \norm{x - y}^2 = 2(\norm{x}^2 + \norm{y}^2).\]
  The inner product is given by
  \[\ab{x, y} = \f 14 \p{\norm{x + y}^2 - \norm{x - y}^2}\]
  for all $x, y \in V$.
\end{theorem}

We now look at a fundamental theorem in linear algebra.
\begin{theorem}[Cauchy-Schwarz Inequality]
  Let $V$ be an inner product space with an induced norm. Then for all $x, y \in V$
  \[\abs{\ab{x, y}} \leq \norm{x}\norm{y}.\]
\end{theorem}
\begin{proof}
  We utilize the discriminant of a quadratic equation and the nonnegativity of the norm. Firstly, if $y = 0$, then the inequality holds trivially. For $y \neq 0$, consider the vector $x + ty$ for some $t \in \bb R$. Then
  \[0 \leq \norm{x + ty}^2 = \ab{x+ty, x+ty} = \norm{y}^2t^2 + 2\ab{x, y}t + \norm{x}^2.\]
  Since the quadratic in $t$ is nonnegative, its discriminant must be nonpositive, giving us
  \[4\ab{x, y}^2 - 4\norm{x}^2\norm{y}^2 \leq 0 \imp \abs{\ab{x, y}} \leq \norm{x}\norm{y}.\]
\end{proof}

\subsection{Eigenvalues and Eigenvectors}
\begin{definition}
  Let $A$ be a square matrix of order $n$. Then, $\la$ is said to be an \emph{eigenvalue} of $A$ if there exists nonzero $\vb x \in \bb R^n$ such that
  \[A\vb x = \la \vb x.\]
  Here, $\vb x$ is called an \emph{eigenvector} corresponding to $\la$.
\end{definition}

To find the eigenvalues of an arbitrary matrix, consider the following:
\[A\vb x = \la \vb x \imp A\vb x - \la \vb x = \vb 0 \imp (A - \la I)\vb x = \vb 0.\]
Since $\vb x$ is nonzero, we require the matrix $A - \la I$ to be singular. That is,
\[\det(A - \la I) = 0.\]
The left side of the equation is a polynomial in $\la$, and is called the \emph{characteristic polynomial}. Therefore the eigenvalues of a matrix $A$ are the \emph{roots} to its characteristic polynomial.

\begin{theorem}
  Eigenvectors corresponding to distinct eigenvalues are linearly independent.
\end{theorem}

\begin{definition}
  A square matrix $A$ is said to be diagonalizable if there exists an invertible matrix $P$ such that
  \[P^{-1}AP = D\]
  where $D$ is a diagonal matrix.
\end{definition}

It turns out that in Definition 1.13, the entries of $D$ along the diagonal are the eigenvalues of $A$, and $P$ contains corresponding eigenvectors in the respective columns of the eigenvalues in $D$.

\subsection{Quadratic Forms}
Certain quadratic functions of several variables may be represented using matrices.
\begin{definition}
  A \emph{quadratic form} is a homogenous polynomial of degree $2$ in multiple variables.
\end{definition}

A quadratic form in $n$ variables may be written with a symmetric matrix $A$ of order $n$ and a vector $\vb x = (x_1, x_2, \dots, x_n)^\top$ of unknowns:
\[Q(x) = \ab{A\vb x, \vb x} = \vb x^\top A\vb x = \sum_{i=1}^n a_{ii}x_{i}^2 + \underset{i \neq j}{\sum_{i=1}^n\sum_{j=1}^n} a_{ij}x_ix_j\]
where $a_{ij}$ is the $(i,j)$th entry of $A$.

\begin{definition}
  Let $A$ be a square matrix of order $n$, and let $p = \ab{A\vb x, \vb x}$. If for all nonzero $\vb x \in \bb R^n$
  \begin{enumerate}
    \item $p > 0$, then $A$ is said to be \emph{positive definite},
    \item $p \geq 0$, \emph{positive semidefinite},
    \item $p < 0$, \emph{negative definite}, and
    \item $p \leq 0$, \emph{negative semidefinite}.
  \end{enumerate}
  Otherwise, $A$ is said to be \emph{indefinite}.
\end{definition}

The definiteness of a matrix may also be realized through its principal minors.

\begin{definition}
  Let $A$ be a square matrix of size $n$. A \emph{principal minor of order $k$} of $A$ may be computed by the following.
  \begin{enumerate}[label=\arabic*.]
    \item Choose any set of $k$ indices $\{i_1, i_2, \dots, i_k\}$ from $\{1, 2, \dots , n\}$.
    \item Keep the rows and columns corresponding to these indices, creating a $k \times k$ submatrix.
    \item Take the determinant of this resulting submatrix to yield a principal minor of order $k$.
  \end{enumerate}
  If the first $k$ indices $\{1, 2, \dots, k\}$ are chosen, then the resulting principal minors called \emph{leading principal minors of order $k$}.
\end{definition}

\begin{theorem}
  A matrix is
  \begin{enumerate}[label=(\alph*), font=\upshape]
    \item positive definite if all leading principle minors are positive,
    \item positive semidefinite if all principal minors are nonnegative,
    \item negative definite if the $k$th leading principal minor has the sign of $(-1)^k$, and
    \item negative semidefinite if all principal minors of order $k$ either have the sign of $(-1)^k$ or are $0$.
  \end{enumerate}
  A matrix that does not fall into the aforementioned categories is indefinite.
\end{theorem}

\end{document}